---
title: "BIOST578 HW1 Xiaowen Tian"
output: html_document
---

```{r}
head(iris)
```
To do a plot of width and length by flower part and species, we need to format data into data frame with columns of length, width, species, and flower part.
First we need to melt data and create id variables.
```{r}
library(reshape2)
iris$Flower_id<-rownames(iris)
irisLong<-melt(iris)
head(irisLong)
```
Then we can split "variable" to get columns of flower_type and mesurement_type.
```{r}
part_measure<-matrix(unlist(strsplit(as.character(irisLong$variable),"\\.")),ncol=2,byrow=T)
irisLong$Flower_part<-part_measure[,1]
irisLong$Measurement<-part_measure[,2]
irisLong<-irisLong[,-3]
head(irisLong)
```
Now, we need to cast data into wide fomat so that it has columns of length and width.
```{r}
irisWide<-dcast(irisLong,Species+Flower_id+Flower_part~Measurement)
head(irisWide)
```
Finally, we can plot the data showing length and width by facetting flower part and flower type.
```{r}
library(ggplot2)
library(ggthemes)
ggplot(data=irisWide,aes(x=Width,y =Length,shape=Flower_part,color = Flower_part)) +
  geom_point()+facet_grid(~Species)+geom_smooth(method = "lm")+ 
  theme_stata()+scale_colour_stata()+ ##add STATA theme 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))
```


