---
title: "BIOST578 HW1 Xiaowen Tian"
output: html_document
---

```{r}
head(iris)
```
To do a plot of width and length by flower part and species, we need to format data into data frame with columns of length, width, species, and flower part.
First we need to melt data and create id variables.
```{r}
library(reshape2)
iris$Flower_id<-rownames(iris)
irisLong<-melt(iris)
head(irisLong)
```
Then we can split "variable" to get columns of flower_type and mesurement_type.
```{r}
part_measure<-matrix(unlist(strsplit(as.character(irisLong$variable),"\\.")),ncol=2,byrow=T)
irisLong$Flower_part<-part_measure[,1]
irisLong$Measurement<-part_measure[,2]
# now we can delete the third column (variable) because that information is captured in the new columns Flower_part and Measurement
irisLong<-irisLong[,-3]
head(irisLong)
```
Now, we need to cast data into wide format so that it has columns of length and width.
```{r}
irisWide<-dcast(irisLong,Species+Flower_id+Flower_part~Measurement)
head(irisWide)
```
Finally, we can plot the data showing length and width by facetting flower part and flower type.
```{r}
library(ggplot2)
library(ggthemes)
ggplot(data=irisWide,aes(x=Width,y =Length,shape=Flower_part,color = Flower_part)) +
  geom_point()+facet_grid(~Species)+geom_smooth(method = "lm")+ 
  theme_stata()+scale_colour_stata()+ ##add STATA theme 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))
```

The plot above has facets just for flower type (or species) and then uses different colors and shapes to represent the different flower parts. If you wanted to produce a plot of `Length` versus `Width` with facets for both flower type *and* flower part, it might look something like this:
```{r}
ggplot(data=irisWide,aes(x=Width,y=Length,shape=Flower_part,color=Flower_part))+
  geom_point()+
  facet_grid(Flower_part~Species)+
  geom_smooth(method='lm')+
  theme_stata()+scale_colour_stata()+ ##add STAT theme
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))
```
